Summary Tables

### 2023-08-23 summary_tables.Rmd
updated use of reference scenario and added comments

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis', warning = F, message = F)

@@

```{r helper_functions, echo = F, message = F}

# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
comprss <- function(tx) { 
  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
        c("","K","M","B","T")[div] )
}

```{r loadLibraries, echo = F, message = F}
suppressWarnings({

@@

```{r load_objects = "asis", echo = F, message = F}
io <- readRDS("results/multi_city/io.rds")
# Assumes that multi_city_script.R has been run till 
# Get names of cities from the io object
cities <<- names(io)[!names(io) %in% 'scen_prop']
LATAM <<- F
AFRICA_INDIA <<- F
NSCEN <- nrow(io$scen_prop)
#cities <- c('accra','sao_paulo','buenos_aires')

# cities <- c('sao_paulo')

@@

# round to
round_to <- 1

# Set plot_mmets to F
plot_mmets <- F

```
# **1st option**
## Scenario definition 
Table displays highest propensity for each distance category, for a given mode.

Table displays highest propensity (%) for each distance category, for a given mode.

In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown.

Distance categories are:

0-2 km = \{trip distance < 2\}

2-6 km = \{2 <= trip distance < 6\}

6+ km = \{6 <= trip distance\}

@@

#rownames(scen_prop) <- paste(rownames(io$scen_prop), "scen", sep = "_")
rownames(scen_prop) <- paste0(substr(toupper(rownames(io$scen_prop)),1,3),"_SC")

kable(scen_prop, headings = "Scenario Proportions")

@@

# ```{r warning=FALSE, echo=F, message=FALSE}
# 
# dat <- list()
# 
# for (city in cities) {
#   
# dat[[city]] <- io[[city]]$trip_scen_sets %>% 
#   mutate(scenario = case_when(
#     scenario == "Scenario 1" ~ "CYC_SC", 
#     scenario == "Scenario 2" ~ "CAR_SC",
#     scenario == "Scenario 3" ~ "BUS_SC",
#     scenario == "Scenario 4" ~ "MOT_SC",
#     scenario == "Baseline" ~ "Baseline")) %>% 
#   group_by(scenario, trip_mode, trip_distance_cat) %>% 
#   summarise(n = dplyr::n(), dist = sum(trip_distance, na.rm = T)) %>% 
#   mutate(city = city)
# 
# td <- dat[[city]]
# 
# for (i in 1:nrow(td)) {
#     val <- td %>% filter(scenario == "Baseline", trip_mode == td$trip_mode[i] & 
#                            trip_distance_cat == td$trip_distance_cat[i] &
#                            city == td$city[i])
#     
#     if (nrow(val) > 0) {
#       td$diff[i] <- val$n - td$n[i]
#       td$trip_prop[i] <- round(td$n[i] / val$n, 2)
#       td$dist_prop[i] <- round(td$dist[i] / val$dist, 2)
#       
#     }
# }
# 
# dat[[city]] <- td
# 
# }
# 
# write_csv(bind_rows(dat), "results/multi_city/scen_prop.csv")
# 
# ```

## Case study specific propensities
```{r message=F, error=F, warning=FALSE}
get_scen_settings <- ithimr::get_scenario_settings()

@@

## Trip proportion
Case study specific trip proportions by mode, for baseline and four scenarios

@@

```{r trip_mode_dist = "asis", echo = F}

@@

for (city in cities) { # Loop for each city
  df <- io[[city]]$trip_scen_sets
  u_trips <- df %>% dplyr::filter(scenario == "baseline") %>% 
    summarise(uid = n_distinct(trip_id)) %>% as.numeric()
  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
    group_by(trip_mode, scenario) %>% 
    summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% 
    spread(key = trip_mode, value = p) %>% 
    mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
  td <- as.data.frame(t(td))
  names(td) <- lapply(append('BASELINE', rownames(scen_prop)), as.character)
  td <- td[-1, ]
  x <- match(row.names(td), trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]

  td1 <- td %>% rownames_to_column()

  names(td1)[1] <- 'stage_mode'

  if (length(trip_prop) == 0){
    trip_prop <- td1 %>% dplyr::select(stage_mode)
  }

  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)

@@

  trip_prop <- inner_join(trip_prop, td1, by = 'stage_mode')

  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
  cat("\n")
} # End loop city

@@

### Baseline trip proportion (%)
```{r}

trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
  janitor::adorn_totals('row')

kable(trip_prop, caption = 'Trip prop (%) for all case studies - only selected modes')

@@

## Distance tables
Case study specific distance tables for baseline and four scenarios

```{r trip_dist = "asis"}

@@

for (city in cities) {

  count_people <- nrow(io[[city]]$synth_pop)
  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
    mutate_if(is.numeric, round, digits = round_to) %>% 
    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  #colnames(td)[3:7] <- rownames(scen_prop)
  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))

  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
  cat("\n")

@@

  if (length(trip_dist) == 0){
    trip_dist <- td1 %>% dplyr::select(stage_mode)
  }

  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)

@@

  trip_dist <- inner_join(trip_dist, td1, by = 'stage_mode')

  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
    mutate_if(is.numeric, round, digits = round_to) %>% 
    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  #colnames(td)[3:7] <- rownames(scen_prop)
  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips)")))

@@

### Baseline average distance (km) per day for all case studies
```{r}

trip_dist <- trip_dist %>% janitor::adorn_totals('row')

kable(trip_dist, caption = "Avg. distance per person (km) per day across all case studies")

@@

## Distance tables scaled by using total population (per year in km)
Case study specific distance tables for baseline and four scenarios

```{r trip_tot_dist = "asis", message=F}

bl_td <- list()

for (city in cities) {

  count_people <- round(sum(io[[city]]$demographic$population))
  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
    mutate_if(is.numeric, list(~round((.) * 365, 3)))
  #colnames(td)[3:7] <- rownames(scen_prop)
  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))

  #

  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  if (length(bl_td) == 0){
    bl_td <- td %>% dplyr::select(stage_mode)
  }

  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))

  td1 <- td %>% dplyr::select(stage_mode, BASELINE)

@@

  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')

  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)

  print(td1)

  # print(kable(td, caption = paste("Distance table (km) for ", city, "( ", sitools::f2si(count_people), " ) per person")))
  cat("\n")

@@

### Baseline total distance for all case studies (per year in km)

@@

# tab <- bl_td %>% gt() %>% fmt_number(columns = 2:ncol(bl_td), decimals = T, suffixing = T)

@@

td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')

for (i in 2:ncol(td)){
  td[, i] <- comprss(td[, i])

print(kable(td, caption = paste("Distance table (km)")))

@@

@@

for (city in cities){

  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), prop = round(n_vals / nrow(.) * 100, 1))
  print(kable(df, caption = paste("Distance by mode by for  ", city)))
}

@@

## True Distance by total population (per day in km)
### Tables

for (city in cities){

  df <- io[[city]]$true_dist

  names(df)[-1] <- lapply(append('BASELINE', rownames(scen_prop)), as.character)

  print(kable(df, caption = paste("True distance by mode per day for  ", city)))
}

@@

<!-- ### Figures -->

<!-- ```{r trip_dist_mode_figs = "asis"} -->

<!-- for (city in cities){ -->

<!--   df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline' & trip_mode %in% c('walking', 'bicycle', 'car', 'motorcycle', 'bus')) %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_mode, trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n()) %>%  -->
<!--     group_by(trip_mode) %>% mutate(n_vals_1 = sum(n_vals)) %>%  -->
<!--     mutate(prop = round(n_vals/n_vals_1 * 100, 1)) -->

<!--   print(ggplot(data=df, aes(x=trip_distance_cat, y=prop)) + -->
<!--           geom_bar(stat="identity", fill="steelblue")+ -->
<!--           facet_wrap(vars(trip_mode)) + -->
<!--           geom_text(aes(label=prop), vjust = -0.3, size = 3.5)+ -->
<!--           theme_minimal() + -->
<!--           theme(panel.spacing = unit(2, "lines")) + -->
<!--           labs(title = paste(city, ' by dist by mode'))) -->
<!-- } -->

<!-- ``` -->

Case study specific duration tables for baseline and four scenarios (per day in minutes)

```{r trip_dur = "asis"}

trip_dur <- list()

l <- list()

for(city in cities){
  count_people <- nrow(io[[city]]$synth_pop)
  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  #colnames(td)[3:7] <- rownames(scen_prop)
  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  l[[city]] <- td

  if (length(trip_dur) == 0){
    trip_dur <- td1 %>% dplyr::select(stage_mode)
  }

  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)

  trip_dur <- inner_join(trip_dur, td1, by = 'stage_mode')

  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (everyone)")))

  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  #colnames(td)[3:7] <- rownames(scen_prop)
  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (people with trips)")))
  cat("\n")
}

### Baseline average duration (mins) per day for all case studies
```{r}

trip_dur <- trip_dur %>% janitor::adorn_totals('row')

kable(trip_dur, caption = "Avg. duration per person (mins) per day across all case studies")

@@

### Change in YLLs per 100,000 {#change_YLL_age_100k}
```{r scen_prop = "asis"}
pop_by_age <- lapply(io[-2], function(x)
  sapply(unique(x$demographic$age), 
         function(y)
           sum(subset(x$demographic, age == y)$population)
         )
  )
yll_rates <- lapply(cities, function(x)
  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000,
        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), 
                                             length = NSCEN)*100000
        )
  )
names(yll_rates) <- cities
round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
```

YLLs per 100,000 people by gender by age group scenario

### Change in YLLs per 100,000 {#change_YLL_gender_age_100k}
```{r scen_prop = "asis"}
#yll_totals <- lapply(io,function(x){
yll_totals <- lapply(io[-2], function(x) {
  temp <- sapply(1:NSCEN,function(y){
    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
    rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
  })
  rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1,
                          function(z)paste0(z[2], '_', z[1]))
  colnames(temp) <- scen_names
  temp
})
yll_rates <- lapply(cities, function(x)
  rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')], 1,
                                    function(z) paste0(z[2],'_',z[1])), 
                              rownames(yll_totals[[x]])),] /
          t(repmat(io[[x]]$demographic$population,NSCEN,1)) * 100000,
        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),
                                             length = NSCEN) * 100000))
names(yll_rates) <- cities
round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
```

>>>>

```{r setup, include=FALSE}

# Script summarises mode splits, distances and durations travelled and looks at health outcomes for the different air pollution, injury and physical activity pathways but also by disease

# WARNING: This script only works if ITHIM-Global has been run in constant mode!

#### It produces the following output documents:

### - html file containing the following information:

# ( - table displaying highest propensity (%) for each distance category, for a given mode. - CURRENTLY commented out as not relevant for BOGOTA, GLOBAL, LATAM or AFRICA_INDIA scenario definitions)

## Mode split
# - for each city a table is displayed showing the Baseline mode split for each distance category

# - for each city a table is given showing the mode split for each scenario

# - one table showing the baseline mode split for all cities

## Distances travelled
# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey by mode for all scenarios (incl Baseline)

# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)

# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities

# - city specific distance tables showing total yearly distances in km travelled by the total  population considered in the model for each mode for all scenarios (incl Baseline)

# - one table giving the city specific total yearly baseline distances in km travelled by the total population for each mode for all cities

# - city specific distance tables showing total daily distances in km travelled by the total population for each mode for all scenarios (incl Baseline)

# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)

# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities

## Travel duration
# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey by mode for all scenarios (incl Baseline)

# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)

# - one table giving the city specific baseline average duration (h) spent travelling by each person in the respective travel surveys for all cities

## Health outcomes - injury, air pollution and physical activity
# - one table for each city showing the change in total YLLs in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age and sex category

# - one table for each city showing the change in total YLLs due to changes in accident fatalities in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in total YLLs due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group

# - one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group

## Health outcomes - disease
# - one table for each city showing the change in deaths in the scenarios compared to the Baseline for each disease

# - one table for each city showing the change in deaths per 100,000 people in the scenarios compared to the Baseline for each disease

# - one table for each city showing the change in YLLs in the scenarios compared to the Baseline for each disease

# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each disease

### - output .csv/.xlsx files containing the following information:

# - the ylls.csv file (results/multi_city/health_impacts/ylls.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people, for each city, disease, scenario, age and sex category. It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and, the age and age and sex specific population levels.

# - the ylls_pathway.csv file (results/multi_city/health_impacts/ylls_pathway.csv)

# - the deaths.csv file (results/multi_city/health_impacts/deaths.csv)

# - the deaths_pathway.csv file (results/multi_city/health_impacts/deaths_pathway.csv)

#### The script performs the following steps:

# - set up helper function to re-format large numbers and re-name scenarios

# ???????????????

# call get_scen_settings.R function to find the Baseline mode splits for each distance category for each city

knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis', warning = F, message = F)

@@

```{r loadLibraries, echo=F, message=FALSE}
suppressWarnings({

@@

```{r helper_functions, echo = F, message = F}
# Function to format numbers with million (M) and billion (B) suffixes
# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
comprss <- function(tx) { 
  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
        c("","K","M","B","T")[div] )
}

# function to rename scenarios
get_qualified_scen_name <- function(cs){
  qualified_scen_name <- ""
  if (cs == 'base' | cs == 'baseline' | cs == 'Baseline')
    qualified_scen_name <- 'Baseline'
  else if(cs == "sc_walk")
    qualified_scen_name <- 'Walking'
  else if(cs == "sc_cycle")
    qualified_scen_name <- 'Bicycling'
  else if(cs == "sc_car")
    qualified_scen_name <- 'Car'
  else if(cs == "sc_motorcycle")
    qualified_scen_name <- 'Motorcycling'
  else if(cs == "sc_bus")
    qualified_scen_name <- 'Public Transport'

  return(qualified_scen_name)
}

# define function to sum the columns of the data, print the data to html and add a caption
sum_and_round_and_print <- function(data, text = '') {
  data <- lapply(data, function(x) round(x, round_to))
  data <- lapply(data, function(x) rbind(x, Total = colSums(x)))
  for (city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

# define function to round the data and print it to html with a caption
round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

```{r load_objects = "asis", echo = F, message = F}

# read in io object which is produced at the end of an ITHIM-Global run
# Assumes that multi_city_script.R has been run 
io <- readRDS("results/multi_city/io.rds")

# Get names of cities from the io object
cities <<- names(io)[!names(io) %in% 'scen_prop']
NSCEN <- nrow(io$scen_prop)

# find reference scenario
ref_scen <- io[[cities[1]]]$ithim_run$reference_scenario
reference_scenario <- get_qualified_scen_name(ref_scen) # update scen name

@@

# define which decimal place to round to
round_to <- 1

```
# **1st option**
## Scenario definition 
<!-- Table displays highest propensity for each distance category, for a given mode. -->

<!-- Table displays highest propensity (%) for each distance category, for a given mode. -->

<!-- In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown. -->

Distance categories are defined as:

0-2 km = \{trip distance < 2km\}

2-6 km = \{2km <= trip distance < 6km\}

6+ km = \{6km <= trip distance\}

@@

rownames(scen_prop) <- paste("sc",rownames(io$scen_prop),  sep = "_")

orig_scen_names_updated <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))

# # rename scenarios
# rownames(scen_prop) <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))

# kable(scen_prop, headings = "Scenario Proportions")

@@

## City mode split for each distance category

```{r message=F, error=F, warning=FALSE}
# call get_scen_settings function to find the mode split for each distance category for each city
get_scen_settings <- ithimr::get_scenario_settings(cities)

@@

## Mode split for each scenario by city
Case study specific trip proportions by mode, for baseline and scenarios

@@

```{r trip_mode_dist = "asis", echo = F}
# find the modal share of for all cities and scenarios

@@

for (city in cities) { # Loop for each city

  # extract the trip data for all scenarios
  df <- io[[city]]$trip_scen_sets

  # find the number of unique trips in the Baseline
  u_trips <- df %>% dplyr::filter(scenario == "baseline") %>% 
    summarise(uid = n_distinct(trip_id)) %>% as.numeric()

  # find the proportion of trips made by each mode in each scenario
  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
    group_by(trip_mode, scenario) %>% 
    summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% 
    spread(key = trip_mode, value = p) %>% 
    mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))

  # find and update scenario names
  scen_names <- td$scenario
  scen_names_updated <- unlist(lapply(scen_names, FUN=get_qualified_scen_name))

  # take transpose
  td <- as.data.frame(t(td))

  # rename columns by using updated scenario names
  names(td) <- scen_names_updated

  # drop first column and re-order by pre-defined mode order
  td <- td[-1, ]
  x <- match(row.names(td), trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]

  # convert row names into column of data and create new dataframe
  td1 <- td %>% rownames_to_column()

  # re-name column
  names(td1)[1] <- 'stage_mode'

  # populate trip_prop list with stage modes
  if (length(trip_prop) == 0){
    trip_prop <- td1 %>% dplyr::select(stage_mode)
  }

  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)

@@

  # add proportions for all cities but only keep stage modes that appear in the trip data of all cities
  trip_prop <- inner_join(trip_prop, td1, by = 'stage_mode')

  # print mode share for individual cities
  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
  cat("\n")
} # End loop city

@@

### Baseline mode split for all cities (%)
```{r}
# create baseline mode share table containing information for all cities

# prep data
trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
  janitor::adorn_totals('row')

# print table to html
kable(trip_prop, caption = 'Baseline mode share (%) for all cities - only selected modes')

@@

## Distance tables
City specific distance tables showing average distance in km travelled by each person by mode for all scenarios (incl Baseline)

```{r trip_dist = "asis"}
# find the average distance a person spends travelling on each mode per day

@@

for (city in cities) { # loop through cities

  # find number of rows / people in the synthetic population
  count_people <- nrow(io[[city]]$synth_pop)

  # extract distances and divide by number of people
  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
    mutate_if(is.numeric, round, digits = round_to) %>% 
    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))

  # update scenario names - cycle, car, bus
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)

  # match with pre-defined mode order
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  # write to .html
  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
  cat("\n")

@@

  # initialise trip_dist
  if (length(trip_dist) == 0){
    trip_dist <- td1 %>% dplyr::select(stage_mode)
  }

  # extract baseline distances
  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)

@@

  # join baseline mode distances for all cities
  trip_dist <- inner_join(trip_dist, td1, by = 'stage_mode')

  ## repeat but only for people who are travelling
  # extract number of people travelling from the trip data rather than the synthetic population
  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
    mutate_if(is.numeric, round, digits = round_to) %>% 
    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips ONLY)")))

@@

### Distance table showing average baseline distances in km travelled by each person by mode for all cities

```{r}
# create table showing average baseline distances by mode for all cities
trip_dist <- trip_dist %>% janitor::adorn_totals('row')

kable(trip_dist, caption = "Avg. baseline distance per person (km) per day across all cities")

@@

## Distance tables scaled by using total model age population (per year in km)

City specific distance tables for baseline and scenarios for the entire population with age range considered in the model

```{r trip_tot_dist = "asis", message=F}
# create distance tables for the yearly distances travelled by the entire population
bl_td <- list()

for (city in cities) { # loop through cities

  # find the total city level population considered in the model only
  #count_people <- round(sum(io[[city]]$demographic$population))

  # extract daily population distances for ages considered in model and multiply by 365 to to get yearly distances
  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
    mutate_if(is.numeric, list(~round((.) * 365, 3)))

  # update scenario names
  colnames(td)[2:ncol(td)] <- scen_names_updated

  # match with pre-defined mode order
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  # initialise bl_td with stage_mode
  if (length(bl_td) == 0){
    bl_td <- td %>% dplyr::select(stage_mode)
  }

  # calculate row and column totals
  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))

  # extract baseline distances
  td1 <- td %>% dplyr::select(stage_mode, Baseline)

@@

  # join baseline distances for all cities
  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')

  # format table
  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)

  # print to html
  cat(paste("Yearly distance (km) travelled by the entire population with ages considered in the model for ", city))
  print(td1)

  cat("\n")

@@

### Baseline total distance for all cities (per year in km)

Baseline distance travelled by entire population with ages considered in the model for all cities

@@

# create baseline yearly distance table for all cities

@@

# add row totals
td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')

# remove all spaces/blanks
for (i in 2:ncol(td)){
  td[, i] <- comprss(td[, i])
}

print(kable(td, caption = paste("Yearly baseline distance table (km) for population with ages considered in the model")))

@@

```{r trip_dist_mode_figs = "asis"}
# for each city create a table showing the total distance, the number of trips and the proportion of trips in each distance category

for (city in cities){ # loop through cities

  # extract trip data for the baseline scenario and calculate the total distance, the number of trips and the proportion of trips in each distance category
  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline' | scenario == 'baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% 
    summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), proportion = round(n_vals / nrow(.) * 100, 1))

 # print to html
  print(kable(df, caption = paste("Distance in travel survey by mode for the ages considered in the model for  ", city)))
}

## Daily population distances (per day in km) for ages considered in model
### Tables

@@

for (city in cities){ # loop through distances

  # extract total daily distances travelled by the entire population with ages considered in the model
  df <- io[[city]]$true_dist

  names(df)[-1] <- scen_names_updated

  print(kable(df, caption = paste("Model age population distance by mode per day for  ", city)))
}

@@

City specific duration tables for baseline and scenarios (per day in minutes)

```{r trip_dur = "asis"}
# create tables showing the average duration each person spends travelling per mode

trip_dur <- list()
l <- list()

for(city in cities){ # loop through cities
  count_people <- nrow(io[[city]]$synth_pop)

  # extract the duration for the synthetic population
  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))

  # update scenario names
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)

  # order by pre-defined mode order
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  # add output to list
  l[[city]] <- td

  # initialise list for all cities
  if (length(trip_dur) == 0){
    trip_dur <- td1 %>% dplyr::select(stage_mode)
  }

  # extract baseline data
  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)

  # create one baseline table containing all city information
  trip_dur <- inner_join(trip_dur, td1, by = 'stage_mode')

  # print to html
  print(kable(td, caption = paste("Duration table (mins) per day for model synthetic population for ", city, "( ", count_people, " ) per person (everyone)")))

  # repeat but only counting people who are actually taking a trip
  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id)) # find unique trip ids rather than rows in synthetic population
  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
  x <- match(td$stage_mode, trip_order$mode)
  x[!is.na(x)]
  td <- td[order(x),]
  row.names(td) <- NULL

  print(kable(td, caption = paste("Duration table (mins) per day based on trips taken in travel survey for ", city, "( ", count_people, " ) per person (people with trips)")))
  cat("\n")
}

@@

### Baseline average duration (mins) per day for all cities

```{r}
# Average baseline duration travelled on each mode for all cities for all people in synthetic population

# calculate the row sum
trip_dur <- trip_dur %>% janitor::adorn_totals('row')

kable(trip_dur, caption = "Avg. duration (min) per person in the synthetic population per day across all cities")

@@

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate change in YLLs compared to the reference scenario per 100,000 people

# extract population statistics
pop_by_age <- lapply(io[-2], function(x)
  sapply(unique(x$demographic$age), 
         function(y)
           sum(subset(x$demographic, age == y)$population)
         )
  )

# calculate ylls per 100,000 people
yll_rates <- lapply(cities, function(x)
  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000, # divide by population
        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), # calculate rate for total YLLs
                                             length = NSCEN)*100000
        )
  )

names(yll_rates) <- cities

# round results and print to html
round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
```

### Change in YLLs compared to reference scenario per 100,000 people by age and sex  {#change_YLL_gender_age_100k}

Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario
```{r scen_prop = "asis"}
cat(' ')
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate changes in YLLs for each scenario compared to the Baseline per 100,000 people by age and sex category

# calculate the total ylls for each scenario summing across all diseases
yll_totals <- lapply(io[-2], function(x) {
  temp <- sapply(1:NSCEN,function(y){
    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
    rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
  })
  rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1,
                          function(z)paste0(z[2], '_', z[1]))
  colnames(temp) <- orig_scen_names_updated
  temp
})
